---
title: "Overview of mvregerr"
author: "P.J Dodd"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Preliminaries

If `devtools` is installed, then the `mvregerr` package can be installed using

```{r}
devtools::install_github('petedodd/mvregerr',dependencies=TRUE)
```

First make sure that the package is loaded before using

```{r}
library(mvregerr)
```

# Simulated data example

## Example 1

TODO without covariates


## Example 2

For these examples, we will need these libraries

```{r}
library(ggplot2)
library(data.table)
```

with covariates

Generate some fake data

```{r}
N <- 2e2   #number of data rows 
K <- 10    #number of outcomes per row
P <- 3     # number of covariates
XX <- matrix(sample(1:2,N*P,replace=TRUE)-1,nrow=N,ncol=P)  #covariates
XX[,1] <- 1                             #intercept terms
BB <- matrix(nrow=P,ncol=K)             #TODO sort out this to have some actual pattern!!
for(k in 1:P) BB[k,] <- 1/k               #some covariates more influential than others
BB[1,] <- 1*(1:K) * (K - (1:K))
YY <- rnorm(n=N*K,mean=c(XX %*% BB),sd=.1)        #
YY <- matrix(YY,nrow=N,ncol=K)                      #back to matrix
eps <- .5         #measurement errors
EPS <- matrix(eps^2,nrow=N,ncol=K)      #errors
ZZ <- YY + matrix(rnorm(N*K,sd=eps),nrow=N,ncol=K)
## fake data done
```

```{r}
## check
ydf <- YY
colnames(ydf) <- letters[1:K]
ydf <- as.data.frame(ydf)
ydf$id <- 1:nrow(ydf)
ydf <- as.data.table(ydf)
ydf <- melt(ydf,id='id')

(plt <- ggplot(data=ydf[id %in%1:10,],aes(x=variable,y=value)) + geom_point())
## (plt <- ggplot(data=ydf,aes(x=variable,y=value)) + stat_summary(fun.data="mean_cl_boot"))

## MLE/MAP attempt for B
xi <- t(XX) %*% XX
## xi <- xi %x% diag(K)
xi <- diag(K) %x% xi                    #TODO THIS WAS WRONG!!
yy <- t(XX) %*% YY  %*% diag(K)
vy <- c(yy)
bmap <- solve(xi,vy)
bmap <- matrix(bmap,nrow=P,ncol=K)
ypred <- XX %*% bmap
colnames(ypred) <- letters[1:ncol(ypred)]
ypdf <- as.data.table(ypred)
ypdf$id <- 1:N
ypdf <- melt(ypdf,id='id')
plt + geom_point(data=ypdf[id %in%1:10,],col=3) + facet_wrap(~id)
```

## ## compare MLE and truth
## yb1 <- copy(ydf)
## yb2 <- copy(ypdf)
## yb1[,true:=value]
## yb2[,MLE:=value]
## ybx <- merge(yb1[,.(id,variable,true)],yb2[,.(id,variable,MLE)],by=c('id','variable'))

## ggplot(data=ybx,aes(x=true,y=MLE,col=variable)) +
##   geom_point() +
##   geom_abline(intercept = 0,slope=1) +
##   coord_fixed()

```{r}
test0 <- mvregerrGS(ZZ,EPS,XX, nchain = 4*50,init=list(Psi=diag(K),nu=5),
                  every = 20,record = c('Y','beta'))

## look at beta

plot(x=(c(BB)),y=c(test0$beta[[1]]))
for(i in 2:50) points(x=(c(BB)),y=c(test0$beta[[i]]))
abline(a=0,b=1,lty=2,col=2)

plot(x=sqrt(c(BB)),y=c(test0$beta[[1]]))
for(i in 2:50) points(x=sqrt(c(BB)),y=c(test0$beta[[i]]))
curve(x^2,from = 0,to=sqrt(max(BB)),col=2,lty=2,add=TRUE)
```

## getYinCnN <- function(X,n) do.call('rbind',lapply(X=X,FUN=function(x)x[n,]))

TODO plot chains
```{r}
y0 <- do.call('rbind',test0$Y)
colnames(y0) <- letters[1:ncol(y0)]
y0df <- as.data.frame(y0)
y0df$id <- 1:nrow(test0$Y[[1]])
y0df$rep <- rep(1:length(test0$Y),each=nrow(test0$Y[[1]]))
y0df <- as.data.table(y0df)
y0df <- melt(y0df,id=c('id','rep'))


ggplot(data=ydf[id%in%1:12,],aes(x=variable,y=value)) + 
  facet_wrap(~id) + 
  geom_point(data=ypdf[id%in%1:12,],col=3,shape=2) +
  geom_violin(data=y0df[id%in%1:12 & rep>20,],col=2)  +
  geom_point(size=2) 

## ## TODO
## ## as predictions!
## test1 <- getGibbs(ZZ,EPS,XX, nchain = 4*50,init=list(Psi=diag(K),nu=5),every = 20,XP=XX)

## y1 <- do.call('rbind',test1$YP)
## colnames(y1) <- letters[1:ncol(y1)]
## y1df <- as.data.frame(y1)
## y1df$id <- 1:nrow(test1$Y[[1]])
## y1df$rep <- rep(1:length(test1$Y),each=nrow(test1$YP[[1]]))
## y1df <- as.data.table(y1df)
## y1df <- melt(y1df,id=c('id','rep'))


## ggplot(data=ydf[id%in%1:12,],aes(x=variable,y=value)) + 
##   facet_wrap(~id) + 
##   geom_violin(data=y1df[id%in%1:12 & rep>20,],col=2)  +
##   geom_point(size=2) 

```